### 2024-10-30 Article/Background_sections/Basic_Notions.tex
Def of subobjects, lemma 1.2.13, 1.2.15

\begin{definition}[Category]\label{def:category}
    A \emph{category} $\cat{C}$ comprises:
    \begin{enumerate}
        \item A collection of \emph{objects} $\Ob(\cat{C})$;
        \item A collection of \emph{arrows} (or \emph{morphisms}) $\Hom(\cat{C})$, often called \emph{homset}.
    \end{enumerate}
    Two operators, $dom$ and $cod$, that map every morphism $f \in \Hom(\cat{C})$ to two objects, respectively, its \emph{domain} and its \emph{codomain}. In case $dom\ f = A$ and $cod\ f = B$, we will write $f: A \rightarrow B$. The collection of morphisms from an object $A$ to an object $B$ is denoted as $\cat{C}(A, B)$.
    An operator $\circ$ of \emph{composition} maps every couple of morphisms $f$, $g$ with $cod\ f = dom \ g$ (in this case $f$ and $g$ are said to be composable) to a morphism $g \circ f : dom\ f \rightarrow cod \ g$. The composition operator is associative, i.e., for each composable arrows $f$, $g$ and $h$, it holds that
    $$
        h \circ (g \circ f) = (h\circ g) \circ f
    $$
    For each object $A$, an \emph{identity} morphism $id_A : A \rightarrow A$ (or, when it is clear from the context, just denoted $A$)  such that, for each $f: A \rightarrow B$:
    \[
        id_B \circ f = f = f \circ id_A 
    \]
\end{definition}

@@

\begin{prop}\label{prop:epi_mono_prop}
    The following statements hold for every pair of composable arrows $f$ and $g$ for any category $\cat C$:
    \begin{enumerate}
        \item if both $f$ and $g$ are mono, then $g \circ f$ is mono;
        \item if $g \circ f$ is mono, then $f$ is mono;
        \item if both $f$ and $g$ are epi, then $g \circ f$ is epi;
        \item if $g \circ f$ is epi, then $g$ is epi.
    \end{enumerate}
\end{prop}
\todo{val la pena osservare la relazione di equivalenza tra sottoggetti corrisponde ad avere un isomorfismo tra i domini}

@@

\begin{proof}
    Before we start proving the statement, we have to note that, given a function $t: A \to B$, it is possible to decompose it as a composition of a injective function and a surjective function, considering the function $A \to t(A)$ sending each element onto its image along $t$, and then applying the inclusion $t(A) \to B$, and such functions are unique. Another way to factorize $t$ is via a composition of a surjective function and an injective function. Consider the equivalence relation $\sim$ defined on $A$, where $a \sim a'$ whenever $t(a) = t(a')$. This equivalence relation induces a map $A \to A/_\sim$, which is surjective. The function $A/_\sim \to B$, mapping each equivalence class onto its image along $t$ is then injective, and this factorization in unique too.

    Let now be $f = f_i \circ f_s$ be the decomposition of $f$ with $f_s$ surjective (i.e., epi in $\Set$) and $f_i$ injective (i.e., mono in $\Set$), and $g = g_i \circ g_s$ be the decomposition of $g$ with $g_i$ injective and $g_s$ surjective, having the following situation.
    \[
        \begin{tikzcd}
            X \ar[r, "{f_s}"] \ar[d, "e"] & f(X) \ar[r, "{f_i}"] & M \ar[d, "m"] \\
            Y \ar[r, "{g_i}"swap] \ar[urr, "h"]& Y/_\sim \ar[r, "{g_s}"swap] & Z
        \end{tikzcd}
    \]
    For the diagram above to commute, must be $f_i \circ f_s = h \circ e$ and $g_s \circ g_i = m \circ h$\todo{completare!!!}
\end{proof}

>>>>

\begin{definition}[Category]\label{def:category}
    A \emph{category} $\cat{C}$ comprises:
    \begin{enumerate}
        \item A collection of \emph{objects} $\Ob(\cat{C})$;
        \item A collection of \emph{arrows} (or \emph{morphisms}) $\Hom(\cat{C})$, often called \emph{homset}.
    \end{enumerate}
    Two operators, $dom$ and $cod$, that map every morphism \-$f \in \Hom(\cat{C})$ to two objects, respectively, its \emph{domain} and its \emph{codomain}. In case $dom\ f = A$ and $cod\ f = B$, we will write $f: A \rightarrow B$. The collection of morphisms from an object $A$ to an object $B$ is denoted as $\cat{C}(A, B)$.
    An operator $\circ$ of \emph{composition} maps every couple of morphisms $f$, $g$ with $cod\ f = dom \ g$ (in this case $f$ and $g$ are said to be composable) to a morphism $g \circ f : dom\ f \rightarrow cod \ g$. The composition operator is associative, i.e., for each composable arrows $f$, $g$ and $h$, it holds that
    $$
        h \circ (g \circ f) = (h\circ g) \circ f
    $$
    For each object $A$, an \emph{identity} morphism $id_A : A \rightarrow A$ (or, when it is clear from the context, just denoted $A$)  such that, for each $f: A \rightarrow B$:
    \[
        id_B \circ f = f = f \circ id_A 
    \]
\end{definition}

@@

\begin{prop}\label{prop:epi_mono_prop}
    The following statements hold for every pair of composable arrows $f$ and $g$ for any category $\cat C$:
    \begin{enumerate}
        \item if both $f$ and $g$ are mono, then $g \circ f$ is mono;
        \item if $g \circ f$ is mono, then $f$ is mono;
        \item if both $f$ and $g$ are epi, then $g \circ f$ is epi;
        \item if $g \circ f$ is epi, then $g$ is epi.
    \end{enumerate}
\end{prop}

\begin{obs}\label{obs:equiv_subobj_implies_isom_of_domains}
    Consider two monomorphisms $m: X \to A$ and $n: Y \to A$, and suppose $m \equiv n$ (in the sense of \Cref{def:subobj}). This corresponds to having a isomorphism between $X$ and $Y$. in fact, since $m \leq n$, there exists s unique $h: Y \to X$ such that $m = n \circ h$, and, since $n \leq m$, there exists a unique $k: X \to Y$ such that $n = m \circ k$. But then
    \[
        \begin{split}
            m &= n \circ h \\
              &= m \circ h \circ k
        \end{split}
        \qquad
        \begin{split}
            n &= m \circ k \\
              &= n \circ h \circ k
        \end{split}
    \]
    Since $m$ is mono, we obtain $id_X = h \circ k$, and, since $n$ is mono, $id_Y = k\circ h$, thus, an isomorphism. 
\end{obs}

@@

{\color{red}\begin{proof}
    Before we start proving the statement, we have to note that, given a function $t: A \to B$, it is possible to decompose it as a composition of a injective function and a surjective function, considering the function $A \to t(A)$ sending each element onto its image along $t$, and then applying the inclusion $t(A) \to B$, and such functions are unique. Another way to factorize $t$ is via a composition of a surjective function and an injective function. Consider the equivalence relation $\sim$ defined on $A$, where $a \sim a'$ whenever $t(a) = t(a')$. This equivalence relation induces a map $A \to A/_\sim$, which is surjective. The function $A/_\sim \to B$, mapping each equivalence class onto its image along $t$ is then injective, and this factorization in unique too.

    Let now be $f = f_i \circ f_s$ be the decomposition of $f$ with $f_s$ surjective (i.e., epi in $\Set$) and $f_i$ injective (i.e., mono in $\Set$), and $g = g_i \circ g_s$ be the decomposition of $g$ with $g_i$ injective and $g_s$ surjective, having the following situation.
    \[
        \begin{tikzcd}
            X \ar[r, "{f_s}"] \ar[d, "e"] & f(X) \ar[r, "{f_i}"] & M \ar[d, "m"] \\
            Y \ar[r, "{g_i}"swap] \ar[urr, "h"]& Y/_\sim \ar[r, "{g_s}"swap] & Z
        \end{tikzcd}
    \]
    For the diagram above to commute, must be $f_i \circ f_s = h \circ e$ and $g_s \circ g_i = m \circ h$\todo{completare!!!}
\end{proof}}

### 2024-10-30 Article/Graph_sections/Eq_Graphs_integration.tex
Def of subobjects, lemma 1.2.13, 1.2.15

    To see that this cone is terminal, let $((E, F, G, a, b, c), \tau_i = (\tau_E^i, \tau_V^i, \tau_C^i)_{i \in \cat I})$ be another cone. By construction, we have an arrow $(\tau_E, \tau_V):(E, F, a, b) \to (A, B, s, t)$ such that
    \[
        \begin{tikzcd}
            & E \arrow[dl, dashed, "{\tau_E}" swap] \arrow[dr, "{\tau_E^i}"] & \\
            A \arrow[rr, "{\pi_E^i}" swap] & & A_i 
        \end{tikzcd}
        %
        \qquad
        %
        \begin{tikzcd}
            & F \arrow[dl, dashed, "{\tau_V}" swap] \arrow[dr, "{\tau_V^i}"] & \\
            B \arrow[rr, "{\pi_V^i}" swap] & & B_i 
        \end{tikzcd}
    \]

>>>>

    To see that this cone is terminal, let $((E, F, G, a, b, c), \tau_i = (\tau_E^i, \tau_V^i, \tau_C^i)_{i \in \cat I})$ be another cone. By construction, we have an arrow $(\tau_E, \tau_V):(E, F, a, b) \to (A, B, s, t)$ such that
    \[
        \begin{tikzcd}[row sep = 26, column sep = 26]
            & E \arrow[dl, dashed, "{\tau_E}" swap] \arrow[dr, "{\tau_E^i}"] & \\
            A \arrow[rr, "{\pi_E^i}" swap] & & A_i 
        \end{tikzcd}
        %
        \qquad
        %
        \begin{tikzcd}[row sep = 26, column sep = 26]
            & F \arrow[dl, dashed, "{\tau_V}" swap] \arrow[dr, "{\tau_V^i}"] & \\
            B \arrow[rr, "{\pi_V^i}" swap] & & B_i 
        \end{tikzcd}
    \]
